---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# analyzer

<!-- badges: start -->
<!-- badges: end -->

The goal of analyzer is to make data analysis easy and accessible to everyone by 
automatically performing the common tasks involved in data analysis. With one 
command this can also create an R notebook (.rmd) file and convert it into 
html or pdf file. The notebooks can also be made interactive.

```{r load_libs}
library(analyzer)
library(ggplot2)
```

## Example - Notebook

This is a basic example which shows you how to solve a common problem. By running 
the following command a notebook can be generated:

```{r example, eval=FALSE}
GenerateReport(dtpath = "mtcars.csv",
               catVars = c("cyl", "vs", "am", "gear"),
               yvar = "vs", model = "binClass",
               output_format = 'html_document',
               title = "Report", tempDir = "~/Documents/temp",
               interactive.plots = FALSE)
```

In the above command we are asking to generate an HTML file (output_format = 'html_document')
for the data **mtcars.csv** stored in the working directory (dtpath = "mtcars.csv").
We defined the variables ("cyl", "vs", "am", "gear") as categorical through the 
parameter 'catVars' and variable "vs" as the dependent variable.

For more details on other parameters, look into the help of **GenerateReport** function.
The generated notebook will have all the relevant code snippents for this particular data 
along with the outputs. The notebook can be used as a reference and modified as required.

## Plots

**analyzer** can also be used to automatically generate plots for all the columns in 
the data. If the data has a dependent (response) variable pass it through the **`r 'yvar'`**
argument and its type through **`r 'yclass'`** argument.

```{r plot_var}
# Simple plot for one variable
p <- plottr(mtcars$mpg)
plot(p$x)
```

```{r plot_df}
# default plots for all the variables in mtcars
p <- plottr(mtcars, yvar = "disp", yclass = "numeric")
plot(p$mpg)
```

The left plot is the univariate plot of 'mpg'. Right plot is bivariate plot with 
the dependent variable 'disp'.

External functions can be used to create custom plots. the function **plottr** can 
take 6 functions for plots of different types of variables. Look into the help of plottr 
for more details. 

```{r plot_custom}
# Define a function for plot for continuous independent and Continuous dependent variables
custom_plot_for_continuous_vars <- function(dat, xname, yname, ...) {
  
  xyplot <- ggplot(dat, aes_string(x=xname, y=yname)) +
    geom_point(alpha = 0.6, color = "#3c4fde") + geom_rug()
  xyplot <- gridExtra::arrangeGrob(xyplot,
                                   top = paste0("New plot of ",
                                                xname, " and ", yname)
  )

  return(xyplot)
  
}
```

The parameters dat, xname, yname and ... must present. Additional arguments can be added.

```{r plot_df2, echo=FALSE}
p <- plottr(mtcars, yvar = "disp", yclass = "numeric", 
            FUN3 = custom_plot_for_continuous_vars)
plot(p$mpg)
```

FUN3 arguemnt is used for generating plots for Continuous independent and Continuous dependent variables.
This will update the right plot. Left (Univariate) plots can also be changeed by using FUN1 
(for Continous) and FUN2 (for Categorical) variables.

```{r plot_custom2}
# Define a function for plot for continuous independent and Continuous dependent variables
custom_plot2 <- function(dat, xname, ...) {
  
  # histogram
  p1 <- ggplot(dat, aes_string(x=xname)) +
    geom_histogram(color="black")

  return(p1)
  
}
```

```{r plot_df3, echo=FALSE}
p <- plottr(mtcars, yvar = "disp", yclass = "numeric",
            FUN1 = custom_plot2, 
            FUN3 = custom_plot_for_continuous_vars)
plot(p$mpg)
```


## Correlation and Association
